### 2024-07-09 Article/Article-english.tex
more text

\medskip\noindent\textbf{Proposition.}\emph{
    $
        \forall \alpha \in \mathbb{Z}^{k \times m}, \beta \in \mathbb{Z}^{m \times n}{:}\ \phi_\alpha \circ \phi_\beta = \phi_{\alpha \beta}
    $.
}

@@

>>>>

\medskip\noindent\textbf{Proposition.}\emph{
    $
        \forall \alpha \in \mathbb{Z}^{k \times m}, \beta \in \mathbb{Z}^{m \times n}{:}\ \phi_\alpha \circ \phi_\beta = \phi_{\alpha \beta}
    $.
}\medskip

@@

Since $\phi_\alpha$ is the~homomorphism, its image $\mathrm{Im}(\phi_\alpha)$ is a~subgroup in~$(K^\times)^n$.
Next we'll study two questions: whether $\mathrm{Im}(\phi_\alpha)$ is an~algebraic subgroup and whether
any algebraic subgroup is represented by~some $\phi_\alpha$.

For this, first of all, we note that the~system of binomial equations $z^{\beta'_i} = z^{\beta''^i}$ is equivalent
to~the system $z^{\beta'_i - \beta''_i} = 1$, that is, exactly the~kernel of the~operator $\phi_{\beta'_i - \beta''_i}$;
therefore, the~first question is equivalent to whether a~given homomorphism $\phi_\alpha$ can be extended
to~the~exact sequence $(K^\times)^k \xrightarrow[]{\phi_\alpha} (K^\times)^n \xrightarrow[]{\phi_\beta} (K^\times)^m$.

\medskip\noindent\textbf{Lemma~5.}\emph{
    Let $G$ be a~group, $H$ be an~abelian group, $f \in \mathrm{Hom}(G, H)$, $g \in \mathrm{Hom}(H, G)$, and~$g \circ f = 1_G$.
    Then the~sequence $G \xrightarrow[]{f} H \xrightarrow[]{f \circ g - 1_H} H$ is exact.
}\medskip

\noindent{\it Proof.}
    Indeed, $(f \circ g - 1_H) \circ f = f \circ g \circ f - f = f - f = 0$, that is, $\mathrm{Im}(f) \subseteq \ker(f \circ g - 1_H)$.
    Conversely, let $h \in \ker(f \circ g - 1_H)$. Then $f(g(h)) - h = 0 \Leftrightarrow h = f(g(h))$, that is, $h \in \mathrm{Im}(f)$
    and $\ker(f \circ g - 1_H) \subseteq \mathrm{Im}(f)$.
\hfill$\Box$\medskip

\medskip\noindent\textbf{Lemma~6.}\emph{
    $\phi_\alpha$ for $\alpha \in \mathbb{Z}^{n \times k}$ can be represented as some image $\mathrm{Im}(\phi_\beta)$
    if the~mappings $g \mapsto g^{\varepsilon_i}$ are surjective in~$G$ for all $\varepsilon_i$ из~$\mathrm{SNF}(\alpha)$.
}\medskip

\noindent{\it Proof.}
    Let us again write $\alpha$ as $\alpha = \beta_1^{-1} \varepsilon \beta_2^{-1}$ where
    $
        \varepsilon = \mathrm{diag}^{n \times k}_r(\varepsilon_1, \ldots, \varepsilon_r).
    $

    Consider $\delta = \mathrm{diag}^{n \times r}_r(1, \ldots, 1)$ и~$\alpha' = \beta^{-1} \delta$.
    $\phi_{\alpha'}$ is injective since it is a~composition of~injective functions. On~the~other hand, it is obvious that:
    $$
        \mathrm{Im}(\phi_\alpha) = \phi_{\beta_1^{-1}} (\phi_\varepsilon (\phi_{\beta_2^{-1}} (G^k)))
                                 = \phi_{\beta_1^{-1}} (\phi_\varepsilon (G^k))
                                 = \phi_{\beta_1^{-1}} (\phi_{\delta} (G^r))
                                 = \mathrm{Im}(\phi_{\alpha'}).
    $$

    It is easy to see that $\beta' = \delta^{\top} \beta$ is a~left inverse of~the~matrix $\alpha'$,
    and hence $\phi_{\beta'}$ is a~left inverse of~$\phi_{\alpha'}$. Since $(\phi_{\alpha'} \circ \phi_{\beta'}) \phi_{-E} = \phi_{\alpha' \beta' - E}$,
    we only need to~apply lemma~5.
\hfill$\Box$\medskip

\medskip\noindent\textbf{Theorem~2.}\emph{
    Any parameterization $\phi_\alpha$ for $\alpha \in \mathbb{Z}^{n \times k}$ defines an~algebraic subgroup
    of~$(K^\times)^n$ if the~field $K$ is algebraically closed.
}\medskip

The~answer to~the~second question is somewhat more complicated. For example, the~equation $z^n = 1$ defines an~algebraic
group for any $n \in \mathbb{Z}$ consisting of~$n$~points on~$\mathbb{C}^\times$; but if $n \neq 0, \pm 1$
it is easy to see that it cannot be parameterized by~any $\phi_\alpha$.

Indeed, notice that the~group $\mathrm{Im}(\phi_\alpha)$ is isomorphic to~the~group $(\mathbb{C}^\times)^r$.
In~the~proof of~lemma~6 we saw that for any $\phi_\alpha$ we can construct an~injective $\phi_{\alpha'} : (\mathbb{C}^\times)^r \rightarrow (\mathbb{C}^\times)^n$
such that $\mathrm{Im}(\phi_\alpha) = \mathrm{Im}(\phi_{\alpha'})$; but then $\phi_{\alpha'}$ defines the~desired isomorphism.
Thus, the~following proposition holds.

\medskip\noindent\textbf{Proposition.}\emph{
    For any matrix $\alpha \in \mathbb{Z}^{n \times k}$ the~group $\mathrm{Im}(\phi_\alpha)$ is isomorphic to an~algebraic torus
    of~dimension at mosk $k$.
}\medskip

Moreover, since the~mapping $\phi_{\alpha'}$ is polynomial, it is, in particular, continuous in~the~standard topology
on~$K = \mathbb{C}$. The~set $(\mathbb{C}^\times)^r$ is connected, and, as we know, the~continuous image
of~a~connected set is connected, so next proposition is also true.

\medskip\noindent\textbf{Proposition.}\emph{
    $\mathrm{Im}(\phi_\alpha) \subseteq (\mathbb{C}^\times)^n$ is connected.
}\medskip

Now notice that the~group $z^n = 1$ is connected exactly when $n = 0, \pm 1$.
This indicates that the~criterion for the~existence of a~monomial parameterization
for~an~algebraic torus subgroup is connectedness.

We denote the~group of~$n$-th root of unity in~the~field $K$ (given by~the~equation $z^n = 1$) by~$\omega_K(n) \subseteq K^\times$.
For~a~vector $x \in \mathbb{Z}^k$ we denote $\omega_K(x) = \omega_K(x_1) \times \ldots \times \omega_K(x_k) \subseteq (K^\times)^k$.
For brevity, we will write $\omega(x) = \omega_\mathbb{C}(x)$. It is known that the~group $\omega(x)$ is isomorphic
to~the~group $\mathbb{Z} / x_1 \mathbb{Z} \times \ldots \times \mathbb{Z} / x_k \mathbb{Z}$.

To~prove the~criterion described before, consider the~number $\Pi(\alpha) = |\omega_K(\varepsilon)|$
where $|G|$ is the~order of~the~group $G$. $|G \times H| = |G| |H|$, so $\Pi(\alpha) = |\omega_K(\varepsilon_1)| \cdot \ldots \cdot |\omega_K(\varepsilon_r)|$.
Since $\omega(n)$ is isomorphic to~$\mathbb{Z} / n \mathbb{Z}$, in~case of~the~field $K = \mathbb{C}$ it is also true that
$\Pi(\alpha) = \varepsilon_1 \cdot \ldots \cdot \varepsilon_r$.

\medskip\noindent\textbf{Lemma~7.}\emph{
    If $\Pi(\beta) = 1$, then there is $\alpha$ such that $\ker(\phi_\beta) = \mathrm{Im}(\phi_\alpha)$.
}\medskip
